A user-extensible and adaptable parser architecture

### 2008-05-30 papers/sgai-2008/paper.tex
Resolved most XXX items.

XXX IN THE INTRODUCTION\@: WHY IS THIS RELEVANT TO AN AI CONFERENCE\@?

XXX MAYBE ADD AN EXAMPLE RULE AND LOG LINE, IF THERE'S SUFFICIENT SPACE\@.

@@

Existing parsers were considered, but their parsing coverage was too low,
their parsing too inexact, and/or they did not extract sufficient data for
the purposes of this project.  The effort required to adapt and improve an
existing parser was judged to be greater than the effort to write one from
scratch, as the techniques used by the existing parsers severely limited
their potential: some ignored the majority of log lines, parsing specific
log lines with great accuracy, but in an inextensible fashion; others
sloppily parsed the majority of log lines, but were incapable of
distinguishing between log lines of the same type, e.g.\ rejections, and
extracted only a portion of the data required for this project.

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejections), accompanied by a multitude of
precise patterns (\textit{rules\/}), each of which matches all log lines of
a specific type, and only that type (e.g.\ rejection by a specific
\DNSBL{}).

This architecture is ideally suited to parsing inputs where the input is
not fully understood or does not follow a fixed grammar: the architecture
warns about unparsed inputs and other errors, but continues on parsing as
best it can, allowing the developer of a new parser to decide which
deficiencies are most important and require attention first, rather than
being forced to fix the first error which arises.

@@

Actions may also return a modified input line which will be parsed as if
read from the input stream, allowing for cascaded parsing XXX REFERENCE\@.
This is a powerful facility allowing a generic rule and action to parse
part of the input, followed by more specific rules and actions parsing the
remainder.

@@

    \item [Partially context-sensitive actions:]  XXX IS THIS THE CORRECT
        TERMINOLOGY\@?  Actions can consult the results (or lack of
        results) of previous actions when executing, providing some context
        sensitivity.  Actions can not inspect past or upcoming inputs,
        however allowing actions can return a modified input line that will
        be parsed as if read from the input stream, allowing for cascaded
        parsing XXX REFERENCE\@.

@@

The parser's coverage of Postfix log files is separated into two parts: log
line coverage and mail coverage.  The first is important because the parser
should handle all input log lines; the second is equally important because
the parser must properly reconstruct every mail if it is to be useful.
Improving the former is less intrusive, as it just requires new rules to be
written; improving the latter is more intrusive as it requires changes to
actions, and it can also be much harder to notice a deficiency.

@@

XXX DO I NEED THIS PARAGRAPH\@?
Parser coverage is divided into two topics in this section: log lines
covered and mails covered.  The former is initially more important, as the
parser must successfully parse every line if it is to be complete, but
subsequently the latter takes precedence because reproducing the path a
mail takes through Postfix is the aim of the parser.  Increasing the
percentage of log lines parsed is relatively simple and non-intrusive:
adding new rules or modifying existing rules is simplified by the
separation of rules, actions and framework.  Improving the logical coverage
is harder, as the actions taken by Postfix must be reconstructed by the
author, and the new sequence of actions integrated into the existing model
without breaking the existing parsing.  Detecting a deficiency in the
parsing algorithm is also significantly harder than detecting unparsed log
lines, as the parser will warn about any unparsed line, whereas discovering
a flaw in the parser requires understanding of the warnings produced and
the mails remaining in the state table.  Rectifying a flaw in the parser
requires an understanding of both the parser and Postfix's log files, and
investigative work to determine the cause of the deficiency, followed by
further examination of the log files in developing a solution.

@@

Developing a new parser based on this architecture follows a similar
process to developing a new parser using any architecture: study the
inputs, figure out what needs to happen for each category of input
(actions), and write a specification identifying each type of input
(rules); the existing framework should be suitable for use with the
majority of parsers.  If the inputs to be parsed are suitable a large
amount of work can be done with very little effort by using \SLCT{} to
identify similar lines in the input and generate \regexes{} to match those
lines; these \regexes{} will form the basis of the initial ruleset.  When
the commonly occurring inputs are parsed by rules and actions, the process
of identifying and rectifying parsing deficiencies can begin, with each
iteration of the development process improving the parser, similar to
Stepwise Refinement~\cite{stepwise-refinement}.

XXX PUT THIS IN THE INTRODUCTION AS WELL\@.  The Postfix log file parser
based on this architecture provides a basis for systems administrators to
monitor the effectiveness of their anti-spam measures and adapt their
defences to combat the new techniques used by those sending spam.  The
parser is a fully usable application, built to address a genuine need,
rather than a proof of concept whose sole purpose is to illustrate a new
idea; it deals with the oddities and difficulties which occur in the real
world, rather than a clean, idealised scenario developed to showcase the
best features of a new approach.

>>>>

@@

Existing parsers were considered, but their parsing coverage was too low,
their parsing too inexact, and/or they did not extract sufficient data for
the purposes of this project.  The effort required to adapt and improve an
existing parser was judged to be greater than the effort to write one from
scratch, as the techniques used by the existing parsers severely limited
their potential: some ignored the majority of log lines, parsing specific
log lines with great accuracy, but in an inextensible fashion; others
sloppily parsed the majority of log lines, but were incapable of
distinguishing between log lines of the same type, e.g.\ rejections, and
extracted only a portion of the data required for this project.  The parser
written for this project is a fully usable application, built to address a
genuine need, rather than a proof of concept whose sole purpose is to
illustrate a new idea; it deals with the oddities and difficulties which
occur in the real world, rather than a clean, idealised scenario developed
to showcase the best features of a new approach.

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejections), accompanied by a multitude of
precise patterns (\textit{rules\/}), each of which matches all log lines of
a specific type, and only that type (e.g.\ rejection by a specific
\DNSBL{}).  This architecture is ideally suited to parsing inputs where the
input is not fully understood or does not follow a fixed grammar: the
architecture warns about unparsed inputs and other errors, but continues on
parsing as best it can, allowing the developer of a new parser to decide
which deficiencies are most important and require attention first, rather
than being forced to fix the first error which arises.

@@

Actions may also return a modified input line which will be parsed as if
read from the input stream, allowing for cascaded
parsing~\cite{cascaded-parsing}.  This is a powerful facility allowing a
generic rule and action to parse part of the input, followed by more
specific rules and actions parsing the remainder.

@@

    \item [Partially context-sensitive actions and cascaded parsing:]
        Actions can consult the results (or lack of results) of previous
        actions when executing, providing some context sensitivity.
        Actions can not inspect past or upcoming inputs, however allowing
        actions can return a modified input line that will be parsed as if
        read from the input stream, allowing for cascaded
        parsing~\cite{cascaded-parsing}.

@@

The parser's coverage of Postfix log files is separated into two parts: log
line coverage and mail coverage.  The former is initially more important,
as the parser must successfully parse every line if it is to be complete,
but subsequently the latter takes precedence because reproducing the path a
mail takes through Postfix is the aim of the parser.  Improving the former
is less intrusive, as it just requires new rules to be written; improving
the latter is more intrusive as it requires changes to actions, and it can
also be much harder to notice a deficiency.

@@

@@

Developing a new parser based on this architecture follows a similar
process to developing a new parser using any architecture: study the
inputs, figure out what needs to happen for each category of input
(actions), and write a specification identifying each type of input
(rules); the existing framework should be suitable for use with the
majority of parsers.  If the inputs to be parsed are suitable a large
amount of work can be done with very little effort by using \SLCT{} to
identify similar lines in the input and generate \regexes{} to match those
lines; these \regexes{} will form the basis of the initial ruleset.  When
the commonly occurring inputs are parsed by rules and actions, the process
of identifying and rectifying parsing deficiencies can begin, with each
iteration of the development process improving the parser, similar to
Stepwise Refinement~\cite{stepwise-refinement}.  This architecture is
ideally suited to parsing inputs where the input is not fully understood or
does not follow a fixed grammar: the architecture warns about unparsed
inputs and other errors, but continues on parsing as best it can, allowing
the developer of a new parser to decide which deficiencies are most
important and require attention first, rather than being forced to fix the
first error which arises.

The Postfix log file parser based on this architecture provides a basis for
systems administrators to monitor the effectiveness of their anti-spam
measures and adapt their defences to combat the new techniques used by
those sending spam.  The parser is a fully usable application, built to
address a genuine need, rather than a proof of concept whose sole purpose
is to illustrate a new idea; it deals with the oddities and difficulties
which occur in the real world, rather than a clean, idealised scenario
developed to showcase the best features of a new approach.
